---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  dpi = 120,
  fig.width = 6,
  fig.height = 4.5,
  out.width = "60%"
)
```

# ggOceanPlots
**Water mass classification, temperature-salinity (TS) and section plots using ggplot2 to aid the analysis of oceanographic data. R package version `r packageVersion("ggOceanPlots")`**

ggOceanPlots turns a tidy data frame of CTD or bottle data into publication-ready figures:

- **`define_water_type()`** — classify each temperature-salinity observation into a water mass.
- **`ts_plot()`** — temperature-salinity diagrams with isopycnals, water-mass polygons and optional marginal density plots.
- **`section_plot()`** — interpolated oceanographic sections or bubble plots along a transect.

These functions originate from [PlotSvalbard](https://github.com/MikkoVihtakari/PlotSvalbard) and are kept up to date here. The package is tuned for the Svalbard region but accepts custom water-mass definitions. For maps of the same data, see the companion package [ggOceanMaps](https://github.com/MikkoVihtakari/ggOceanMaps).

> **Using an AI assistant?** See [AGENTS.md](AGENTS.md) for a task-oriented guide with the common pitfalls spelled out.

## Installation

The package can be installed from GitHub using [**devtools**](https://cran.r-project.org/package=devtools) (or [remotes](https://cran.r-project.org/package=remotes)):

```{r eval = FALSE}
devtools::install_github("MikkoVihtakari/ggOceanPlots")
```

## Temperature-salinity plots

`ts_plot()` returns a ggplot2 object. By default points are coloured by water type, isopycnals are drawn, and the relevant water-mass polygons are labelled.

```{r ts-basic}
library(ggOceanPlots)

ts_plot(ctd_kongsfjord)
```

The `color` argument is overloaded: `"watertype"` (default) colours by classified water mass, a **column name** scales colour to that variable, and a literal colour name (e.g. `"red"`) paints every point one colour. Marginal density plots can be added with `margin_distr = TRUE`.

```{r ts-area}
ts_plot(ctd_kongsfjord, color = "area", margin_distr = TRUE)
```

Custom or alternative water masses are passed through the `WM` argument (`rijpfjord_watermasses` and `isfjord_watermasses` ship with the package). Set `WM = NULL` to drop the polygons entirely.

```{r ts-rijp}
ts_plot(ctd_rijpfjord, WM = rijpfjord_watermasses)
```

## Water-mass classification

`define_water_type()` returns the classification on its own, or bound to the data:

```{r classify}
head(define_water_type(ctd_kongsfjord, bind = TRUE))
```

## Section plots

`section_plot()` produces an interpolated section (`interpolate = TRUE`) or a bubble plot (the default). A logarithmic y-axis helps when station depths differ greatly, and contour lines can be labelled.

```{r section}
section_plot(ctd_rijpfjord, x = "dist", y = "pressure", z = "temp",
             bottom = "bdepth", interpolate = TRUE, log_y = TRUE,
             contour = c(-1.8, 0, 1, 3))
```

Bubble plots suit sparse samples such as water-bottle data:

```{r bubble}
section_plot(chlorophyll[grepl("KpN.|Kb[0-4]", chlorophyll$Station), ],
             x = "lon", y = "From", z = "Chla", zlab = "Chl a")
```

## Citation

```{r citation, eval = FALSE}
citation("ggOceanPlots")
```

Bug reports and feature requests are welcome at <https://github.com/MikkoVihtakari/ggOceanPlots/issues>.
