---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE, 
  warning = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# ggOceanPlots
**Water mass classification, TS- and section plots using ggplot2 to aid analysis of oceanographic data. R package version `r packageVersion("ggOceanPlots")`**

This package contains water mass classification, temperature-salinity-, and section plot functions from [PlotSvalbard](https://github.com/MikkoVihtakari/PlotSvalbard). These functions are more up-to-date in this package, which may eventually be submitted to CRAN.

Note that the [spatial interpolation function](https://github.com/MikkoVihtakari/ggOceanPlots/blob/main/R/interpolate_spatial.R) is a direct translation from PlotSvalbard and works suboptimally at the moment. Use PlotSvalbard or ggOceanMaps and [the ggplot2 interpolation](https://ggplot2.tidyverse.org/reference/geom_contour.html) instead of ggOceanPlots for spatial interpolation. Revision of the spatial interpolation functions in this package may happen eventually. 

## Installation

The package can be installed from GitHub using the [**devtools**](https://cran.r-project.org/web/packages/devtools/index.html) (or [remotes](https://cran.r-project.org/web/packages/remotes/index.html)) package:

```{r eval = FALSE}
devtools::install_github("MikkoVihtakari/ggOceanPlots")
```