---
title: "Getting started with ggOceanPlots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with ggOceanPlots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.width = 6,
  fig.height = 4.5,
  dpi = 96
)
```

```{r setup}
library(ggOceanPlots)
```

ggOceanPlots helps you go from a tidy data frame of CTD or bottle data to three
kinds of oceanographic figure: temperature-salinity (TS) diagrams, water-mass
classifications, and section plots. It is built on
[ggplot2](https://ggplot2.tidyverse.org/), so the plots can be extended with
ordinary ggplot2 syntax.

## Input data

The functions expect a **long/tidy data frame** with one row per observation.
Column names are supplied as character strings, so your columns can be named
anything. The package ships with example data:

```{r data}
str(ctd_kongsfjord)
```

If you rely on the built-in water masses, use **potential temperature** and
**practical salinity** — the water-mass boundaries and isopycnals assume those
units.

## Temperature-salinity diagrams

The default `ts_plot()` colours points by water type, draws isopycnals, and
labels the relevant water-mass polygons.

```{r ts-default}
ts_plot(ctd_kongsfjord)
```

### Controlling colour

The `color` argument behaves differently depending on what you give it:

```{r ts-color}
# Scale colour to a column in the data
ts_plot(ctd_kongsfjord, color = "area")

# A single fixed colour (any name from colors())
ts_plot(ctd_kongsfjord, color = "red")
```

### Marginal distributions and zoom

`margin_distr = TRUE` adds kernel-density plots to the axes. Note that with
marginal plots `ts_plot()` draws directly to the graphics device and does **not**
return a ggplot object. `zoom = FALSE` shows the full water-mass diagram rather
than zooming to the data.

```{r ts-extras, fig.keep = "last"}
ts_plot(ctd_kongsfjord, zoom = FALSE)
```

## Classifying water masses

`define_water_type()` returns the classification, either as a factor or bound to
the data frame:

```{r classify}
table(define_water_type(ctd_kongsfjord))
```

### Custom water masses

A water-mass (`WM`) data frame defines rectangles in salinity (x) – temperature
(y) space. Overlaps are resolved by row order: the first matching row wins, so
each point gets exactly one type. Use `Inf` / `-Inf` for open-ended bounds.

```{r custom-wm}
my_wm <- data.frame(
  type     = c("Warm", "Cold"),
  abb      = c("W", "C"),
  sal.min  = c(34.5, -Inf), sal.max = c(Inf, 34.5),
  temp.min = c(2, -Inf),    temp.max = c(Inf, 2)
)

ts_plot(ctd_kongsfjord, WM = my_wm)
```

The package also bundles `rijpfjord_watermasses` and `isfjord_watermasses`.

## Section plots

`section_plot()` interpolates a section (`interpolate = TRUE`) or draws a bubble
plot (`interpolate = FALSE`, the default). Supply the x (distance), y (depth) and
z (variable) columns by name.

```{r section}
section_plot(ctd_rijpfjord, x = "dist", y = "pressure", z = "temp",
             bottom = "bdepth", interpolate = TRUE)
```

A logarithmic y-axis reveals near-surface structure when stations differ greatly
in depth. With `log_y = TRUE`, supply `ybreaks` in real depths — the transform is
applied internally.

```{r section-log}
section_plot(ctd_rijpfjord, x = "dist", y = "pressure", z = "temp",
             bottom = "bdepth", interpolate = TRUE, log_y = TRUE,
             contour = c(-1.8, 0, 1, 3))
```

For sparse data such as water-bottle samples, a bubble plot is often clearer:

```{r bubble}
section_plot(chlorophyll[grepl("KpN.|Kb[0-4]", chlorophyll$Station), ],
             x = "lon", y = "From", z = "Chla", zlab = "Chl a")
```

The interpolation engine (`interpolate_section()`) supports `"mba"` (default,
multilevel B-spline; needs a square grid so `nx == ny`), `"linear"`, `"spline"`
and `"idw"`.

## Extending and saving

`ts_plot()` (without `margin_distr`) and `section_plot()` return ggplot2 objects,
so the usual tools apply:

```{r save, eval = FALSE}
p <- ts_plot(ctd_kongsfjord) + ggplot2::ggtitle("Kongsfjorden")
ggplot2::ggsave("ts.png", p, width = 6, height = 5, dpi = 300)
```

## See also

- [ggOceanMaps](https://mikkovihtakari.github.io/ggOceanMaps/) for plotting the
  same data on bathymetric maps. For spatial interpolation, prefer ggOceanMaps
  over the legacy `interpolate_spatial()` in this package.
